Leadership across cultures and contexts: an exploratory analysis

### 2020-05-14 leadership_across_cultures_contexts.Rmd
More work on fivefold section

## The six-fold way

In our view, the dominance-prestige and "dual-process" models of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers". In short, these models have conflated "benefits" with knowledge and expertise, and "costs" with physical formidability. Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability. The "dual" in dual process comes from the fact that one individual can influence others either by providing benefits or by imposing costs. Moreover, not all leaders have status (e.g., managers), and not all high status individuals are leaders. We therefore tenatively advance an alternative scheme whereby leaders possess some combination of knowledge/expertise and physical formidability which they use to provide benefits or impose costs. If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status. Hence, we should be able to find examples of leaders with various combinations of these five elements.

In our data, we combinded a subset of our leader dimensions into these five elements. We then calculated the fraction of text records with evidence for one or more of the five elements. See Figure \@ref(fig:fivefold).

```{r fivefold, fig.width=10, fig.cap=""}
upset(df_fivefold, order.by='freq', nintersects = 19, text.scale = 2)
```

>>>>

## The five-fold way

In our view, the dominance-prestige and "dual-model" models of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers". In short, these models have conflated "benefits" with knowledge and expertise, and "costs" with physical formidability. @van_vugt_dual_2019, for instance, state that the source of deference for prestige-style leadership is "Information asymmetry", and for dominance-style, "Power asymmetry" (their Table 1). Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006].

We therefore tenatively advance an alternative scheme whereby individuals possess some combination of knowledge/expertise and physical formidability which they use to provide benefits and impose costs. If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016]. Hence, we should be able to find examples of leaders with various combinations of these five elements.

In our data, we used the high status quality as "High status", the prosociality feature as "Provide benefits", any one of punishment, aggressive, feared, and killer as "Impose costs", knowledge/intellect and/or experience/accomplished as "Knowledge", and physically formidible and/or physically healthy as "Physical". We then calculated the fraction of text records with evidence for one or more of the five elements. See Figure \@ref(fig:fivefold).

```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic dimensions of leadership among all text records (an 'upset' plot) [@gehlenborg2019]."}
upset(df_fivefold, order.by='freq', nintersects = 19, text.scale = 2)
```
